### 2019-05-31 talk/animation/spaces12.tex

\documentclass[tikz,border=3.14pt]{standalone}
\usepackage{../mymacros}
\usepackage{tikz-3dplot}
\usetikzlibrary{fadings}

\tikzfading[name=fade out,
inner color=transparent!0, outer color=transparent!100]

\begin{document}
\begin{tikzpicture}[]
\draw[] (0,0,0) ellipse (2 and .75) node[right, shift={(1,1)}]{$\Theta$};
\draw [fill, color=blue!60!white] (-0.7,-0.3,0) circle (1.5pt) node[above](a){$\vtheta_A$};
\draw [fill, color=red!60!white] (0.9,0.1,0) circle (1.5pt) node[above](b){$\vtheta_B$};

\draw[dashed, color=blue!60!white, opacity=0.3] (-1.93,-3.5,0) -- (-0.7,-4.75,0);
\draw[dashed, color=blue!60!white, opacity=0.3] (0.5,-3.5,0) -- (-0.7,-4.75,0);
\draw[dashed, color=red!60!white, opacity=0.3] (-0.33,-3.5,0) -- (0.9,-4.75,0);
\draw[dashed, color=red!60!white, opacity=0.3] (2.1,-3.5,0) -- (0.9,-4.75,0);

\draw[] (0,-3.5,0) ellipse (2.5 and .75) node[right, shift={(1,1)}]{$\mathcal{T}$};
\draw[fill, inner color=blue!60!white, outer color=white, draw=none, color=white] (-0.7,-3.5,0) ellipse (1.25 and .4) node[above, shift={(-.5,.2)}, color=blue!60!white]{$p_{\vtheta_A}$};
%\draw[dashed, color=blue!60!white] (a) -- (-0.7,-3.5,0);
\draw[color=blue!60!white] (-0.7,-5,0) node (c) {$J(\vtheta_A)$};
%\draw[->, dashed, color=blue!60!white] (-0.7,-3.5,0) -- (c);

\draw[] (0,-3.5,0) ellipse (2.5 and .75) node[right, shift={(1,1)}]{$\mathcal{T}$};
\draw[fill, color=red!60!white, path fading=fade out, draw=none] (0.9,-3.5,0) ellipse (1.25 and .4);
\draw[color=red!60!white] (1.4,-3.05,0) node{$p_{\vtheta_B}$};
%\draw[dashed, color=red!60!white] (b) -- (0.9,-3.5,0);
\draw[color=red!60!white] (0.9,-5,0) node (d) {$J(\vtheta_B)$};
%\draw[->, dashed, color=red!60!white] (0.9,-3.5,0) -- (d);

\draw[dashed, color=blue!60!white, opacity=0.3] (-0.7,-0.3,0) -- (-1.93,-3.5,0);
\draw[dashed, color=blue!60!white, opacity=0.3] (-0.7,-0.3,0) -- (0.5,-3.5,0);

\draw[dashed, color=red!60!white, opacity=0.3] (0.9,0.1,0) -- (-0.33,-3.5,0);
\draw[dashed, color=red!60!white, opacity=0.3] (0.9,0.1,0) -- (2.1,-3.5,0);

\draw[line width = 1.5pt, color=green!70!black,-latex] (-0.7,-3.4,0) arc[start angle=180,end angle=0, x radius = 0.8 cm , y radius =1 cm] node[shift={(-0.75,1.2)}]{\bf IS};

%\draw[color=blue!60!white] (-0.7,-5,0) node (c) {$J(\vtheta_1)$};
%\draw[->, dashed, color=blue!60!white] (a) -- (c);
%\draw[color=red!60!white] (0.9,-5,0) node (d) {$J(\vtheta_2)$};
%\draw[->, dashed, color=red!60!white] (b) -- (d);
%\draw[fill=black] (-1.5,-1.5,0) rectangle (1.7,-3.5,0) node[pos=.5, %color=white]{\bf\Large MAB};
\end{tikzpicture}
\end{document}