### 2019-05-31 talk/animation/spaces9.tex

\documentclass[tikz,border=3.14pt]{standalone}
\usepackage{../mymacros}
\usepackage{tikz-3dplot}
\usetikzlibrary{fadings}

\tikzfading[name=fade out,
inner color=transparent!0, outer color=transparent!100]

\begin{document}
\begin{tikzpicture}[]
\draw[] (0,-3.5,0) ellipse (2.5 and .75) node[right, shift={(1.5,1)}]{$\mathcal{T}$};
\draw[fill,color=violet!60!white, path fading = fade out, draw=none] (-0.5,-3.5,0) ellipse (.8 and .4);
\draw[color=violet!60!white] (-0.9,-3.2,0) node{$\Phi_t$};

\draw[fill,color=black!60!white, path fading = fade out, draw=none] (1.,-3.7,0) ellipse (1 and .3);
\draw[color=black!60!white] (1.2,-3.4,0) node{$p_{\vtheta}$};

\draw[line width = 1.5pt, color=orange!90!black,latex-latex] (-0.5,-3.5,0) --  (1.,-3.7,0) node[shift={(-0.75,0.5)}]{$d_2$};

\end{tikzpicture}
\end{document}